### 2019-09-25 revised_dmtcs.d/symmetry_properties.tex
ssr: changes suggested by reviewer

\begin{corollary}\label{cor:ncf_r_symmetric_layers}
(i) An NCF with a default-normalized nested canalyzing representation consisting
of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
(ii) An $r$-symmetric NCF has a default-normalized nested canalyzing representation
with at most $r-1$ layers, and has an unnormalized nested canalyzing
representation with at most $r$ layers.   \QED
\end{corollary}

@@

Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric 
Boolean function of $n$ variables.
We assume that the symmetry groups 
of $f$, denoted by $g_1$, $g_2$, $\ldots$, $g_r$,
are given.
Let $m_i = |g_i|$, $1 \leq i \leq m$.
Thus, in group $g_i$, the number
of variables which can take on the value 1 varies from 0
to $m_i$, $1 \leq i \leq r$.
We also assume that $f$ is
given as a table $T$ of the following form:
each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
where $c_i$ is the number of variables of group $g_i$ which
have the value 1, along with the \{0,1\} value of $f$ for that $r$-tuple.
Thus, $\mu$, the number of rows in $T$, 
is given by  $\mu = \prod_{i=1}^r (m_i+1)$.
As will be seen, our algorithm for determining whether $f$ is
an NCF runs in $O(r\mu)$ time.
The algorithm relies on the following observation.

@@

To estimate the running time, we note that in iteration $j$, we can
determine whether there is a canalyzing variable in $X_j$
in $O(r_j \mu_j)$ = $O(r \mu_j)$ time using 
Observation~\ref{obs:rsym_canalyzing}.
(This is done by considering each variable in $X_j$.)
Once a canalyzing variable is identified, the other steps in that iteration
can be completed in $O(\mu_j)$ time.
Thus, the time for iteration $j$ is $O(r \mu_j)$.
We note that in each iteration, the number of rows in the table
is reduced by a factor of at least 2; that is, $\mu_{j+1} \leq \mu_j/2$.
To see why, suppose the canalyzing variable $x_i$ found in iteration $j$
is in group $g_p$ with $m_p$ variables. 
Thus, in $T_j$, there are $m_p+1 \geq 2$ possibilities for the number of 1's 
in group $g_p$. 
As indicated in Substep~4 of Step~II above, $T_{j+1}$ 
contains only those rows of $T_j$ where all the variables 
in $g_p$ have the same value.
In other words, there is only one possibility for the number of
1's in group $g_p$.
Thus, $\mu_{j+1}$, the number of rows in $T_{j+1}$,
is at most $\mu_j/(m_p+1)$ $\leq$ $\mu_j/2$.
It follows by simple induction that $\mu_j \leq \mu/2^{j-1}$.
Thus, the running time over all the $r$ iterations is given by
$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$, since the
geometric sum is bounded by $2\mu$.
The following theorem summarizes this result.

\begin{theorem}\label{thm:rsym_canalyzing}
The problem of testing whether a given 
$r$-symmetric function $f$ is an NCF can 
be solved efficiently. \QED
\end{theorem}

@@

\hspace*{0.5in} $x_1 \vee (\overline{x}_2 \wedge ( x_3 \vee (\overline{x}_4 \wedge (\cdots ))$

@@

\noindent
For instance,
\begin{align*}
f_6 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge 
      (x_5 \vee \overline{x}_6)) \\
f_7 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 
          \vee (\overline{x}_6 \wedge  x_7)).
\end{align*}
Function $f_n(x_1, x_2, \ldots, x_n)$ is the NCF corresponding to
the following NCF representation:

>>>>

\begin{corollary}\label{cor:ncf_r_symmetric_layers}
(i) An NCF with a default-normalized nested canalyzing representation consisting
of $q$ layers is $2q$-symmetric, and is not $(q-1)$-symmetric.
(ii) An $r$-symmetric NCF has a default-normalized 
nested canalyzing representation with at most $r$ layers. \QED
\end{corollary}

@@

Let $f(x_1, x_2, \ldots, x_n)$ be an $r$-symmetric 
Boolean function of $n$ variables.
We assume that the symmetry groups 
of $f$, denoted by $g_1$, $g_2$, $\ldots$, $g_r$,
are given.
Let $m_i = |g_i|$, $1 \leq i \leq r$.
Thus, in group $g_i$, the number
of variables which can take on the value 1 varies from 0
to $m_i$, $1 \leq i \leq r$.
We also assume that $f$ is
given as a table $T$ of the following form:
each row of $T$ specifies an $r$-tuple $(c_1, c_2, \ldots, c_r)$,
where $c_i$ is the number of variables of group $g_i$ which
have the value 1, along with the \{0,1\} value of $f$ for that $r$-tuple.
Thus, $\mu$, the number of rows in $T$, 
is given by  $\mu = \prod_{i=1}^r (m_i+1)$.
Since there are $\mu$ rows and each row is an $r$-tuple,
the size of the table $T$ is $O(r\mu)$.
As will be seen, our algorithm for determining whether $f$ is
an NCF runs in $O(r\mu)$ time.
The algorithm relies on the following observation.

@@

To estimate the running time, we note that in iteration $j$, we can
determine whether there is a canalyzing variable in $X_j$
in $O(r_j \mu_j)$ = $O(r \mu_j)$ time using 
Observation~\ref{obs:rsym_canalyzing}.
(This is done by considering each variable in $X_j$.)
Once a canalyzing variable is identified, the other steps in that iteration
can be completed in $O(\mu_j)$ time.
Thus, the time for iteration $j$ is $O(r \mu_j)$.
We note that in each iteration, the number of rows in the table
is reduced by a factor of at least 2; that is, $\mu_{j+1} \leq \mu_j/2$.
To see why, suppose the canalyzing variable $x_i$ found in iteration $j$
is in group $g_p$ with $m_p$ variables. 
Thus, in $T_j$, there are $m_p+1 \geq 2$ possibilities for the number of 1's 
in group $g_p$. 
As indicated in Substep~4 of Step~II above, $T_{j+1}$ 
contains only those rows of $T_j$ where all the variables 
in $g_p$ have the same value.
In other words, there is only one possibility for the number of
1's in group $g_p$.
Thus, $\mu_{j+1}$, the number of rows in $T_{j+1}$,
is at most $\mu_j/(m_p+1)$ $\leq$ $\mu_j/2$.
It follows by simple induction that $\mu_j \leq \mu/2^{j-1}$.
Thus, the running time over all the $r$ iterations is given by
$O(r [\sum_{j=1}^{r}(\mu/2^{j-1})])$ = $O(r\mu)$, since the
geometric sum is bounded by $2\mu$.
Recall that the size of input table $T$ 
(which specifies the $r$-symmetric function $f$)
is $O(r\mu)$. 
Thus, the running time of the algorithm
is linear in the size of the input.
The following theorem summarizes this result.

\begin{theorem}\label{thm:rsym_canalyzing}
Suppose an $r$-symmetric function $f$ is given as a table where
each row has an $r$-tuple of the form $(c_1, c_2, \ldots, c_r)$,
with $c_i$ being the number of variables in group $g_i$ that have value 1,
along with the \{0,1\} value of the function for that row.
The problem of testing whether $f$ 
is an NCF can be solved in time that 
is linear in the size of the input. \QED
\end{theorem}

When an $r$-symmetric function $f$ with $n$ variables is
specified using the table representation described in the
statement of Theorem~\ref{thm:rsym_canalyzing},
the size of the table is $O(n^r)$.
When $r$ is \emph{fixed}, the size of 
this representation and the running time 
of the above algorithm are both polynomial functions of $n$.

@@

\hspace*{0.5in} $x_1 \vee (\overline{x}_2 \wedge 
    ( x_3 \vee (\overline{x}_4 \wedge (\cdots ))))$

@@

\noindent
For instance,
\begin{align*}
f_6 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge 
      (x_5 \vee \overline{x}_6))))~~~ \mathrm{and}\\
f_7 &~=~  x_1 \vee (\overline{x}_2 \wedge (x_3 \vee (\overline{x}_4 \wedge (x_5 
          \vee (\overline{x}_6 \wedge  x_7))))).
\end{align*}
Function $f_n(x_1, x_2, \ldots, x_n)$ is the NCF corresponding to
the following NCF representation:
